---
title: "Lecture 1 - issues raised by missing data, a systematic approach, and missingness mechanisms"
subtitle: "Multiple imputation for missing data"
author: "[Jonathan Bartlett (thestatsgeek.com)](https://thestatsgeek.com)"
date: "Oslo, November 2019"
output:
  beamer_presentation:
    keep_tex: true
    slide_level: 2
    toc: true
urlcolor: blue
bibliography: references.bib
link-citations: true
linkcolor: blue
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(pander)

def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

## Course aims

- Understand the effects of missing data on statistical analyses
- Learn about the assumptions under which simple methods for handling missing data are valid
- Learn about principled statistical methods for handling missing data, specifically multiple imputation


# Missing data - what's the big deal and a systematic approach

## Why is this necessary?

- Missing data commonly arise in empirical research.
- They cause a loss of information, and arguably more importantly, may introduce bias into inferences.
- They are often inadequately handled in both observational and experimental studies.
- For example, [@karahalios2012review] reviewed the reporting and handling of missing data in longitudinal measurements in cohort studies.
- They found that reporting of missing data was inconsistent and inappropriate statistical methods continue to be used (in this field at least).
- Scientific journals and bodies increasingly recognise the importance of careful handling of missing data.


## Missing data in trials - the problem and its prevention

- A US National Research Council (NRC) report was recently published on the prevention and treatment of missing data in trials [@NRC2010;@Little2012].
- They noted that missing data have seriously compromised inferences from clinical trials in the past.
- They concluded that the assumption that analysis methods can compensate for missing data are not justified.
- The panel therefore recommended strategies for minimizing missing data in trials.

## Missing data in trials - six recommended principles (steps)
Based on [@Little2012]

1. Find out if values are missing are relevant for the intended analysis.
2. Formulate a well defined causal primary measure of treatment effect.
3. Document and investigate the reasons for missing data.
4. Decide on a primary set of assumptions about missing data.
5. Perform an analysis using a statistical method which is valid under the assumption chosen in 4.
6. Perform a sensitivity analysis to explore robustness to plausible deviations from the assumption in 4.

## A principled approach

- We will attempt to follow such an approach.
- Thinking more generally, outside of clinical trials, step 2. consists of specifying our substantive model or quantitiy of interest.


## Example

- e.g. consider the following break down of smoking status (for males in THIN from [@Marston2010].
- Our objective is to estimate the marginal distribution of smoking status in the population.
	
Smoking status  | n (% of sample) | (% of those observed)
--------------- | --------------- | ----------------------
Non             | 82,479 (36)     | (48)
Ex              | 30,294 (13)     | (18)
Current         | 57,599 (25)     | (34) 
Missing         | 56,661 (25)     | n/a 

- Are the %s in the last column unbiased estimates?

# Missingness mechanisms


## Rubin's classification

- Our first step is to think about the mechanism causing a variable (e.g. smoking status) to be missing.
- Rubin developed a classification for missing data `mechanisms' [@Rubin:1976].
- We introduce the three types in a very simple setting.
- We assume we have one fully observed variable $Y_{1}$ (age), and one partially observed variable $Y_{2}$ (blood pressure (BP)).
- We will let $R$ indicate whether $Y_{2}$ is observed ($R=1$) or is missing ($R=0$).
- Note $Y_{2}$ is not necessarily the `outcome' in our final analysis.

## Missing completely at random

- The missing values in BP ($Y_{2}$) are said to be missing completely at random (MCAR) if missingness is independent of BP ($Y_{2}$) and age ($Y_{1}$).
- i.e. those subjects with missing BP do not differ systematically (in terms of BP or age) to those with BP observed.
- In terms of the missingness indicator $R$, MCAR means

$$P(R=1|Y_{1},Y_{2})=P(R=1)$$

## Example - blood pressure (simulated data)

To illustrate, we consider some simulated data on age (categorised) and systolic blood pressure.
```{r echo=FALSE}
expit <- function(x) {exp(x)/(1+exp(x))}
set.seed(1235)
n <- 200
ageCat <- factor(c(rep(0,n/2),rep(1,n/2)), labels=c("30-50 years", "50-70 years"))
bp <- 110+30*(as.numeric(ageCat)-1)+rnorm(n,sd=20)
r <- 1*(runif(n)<expit(as.numeric(ageCat)-1))
rCat <- factor(r, labels=c("BP missing", "BP observed"))
bpFull <- data.frame(ageCat,bp,rCat)
bp[r==0] <- NA
bpObs <- data.frame(ageCat,bp,rCat)
```

```{r, echo=FALSE}
library(foreign)
write.dta(bpObs, "bpObs.dta")
```

```{r, size="small"}
summary(bpObs)
```

## Checking MCAR
- With the observed data, we could investigate whether age $Y_{1}$ is associated with missingness of blood presure ($R$).
- If it is, we can conclude the data are \textcolor{red}{not} MCAR.
- If it is not, the data are consistent with MCAR, although it is still possible that it is MNAR.
- It is possible (though arguably unlikely in this case) that BP is associated with missingness in BP, even if age is not.

## Checking MCAR

To examine whether BP is plausibly MCAR, we compare the proportion of missingness between the two age categories:

```{r, echo=FALSE, size="small"}
pander(addmargins(table(bpObs$ageCat, rCat)))
```

## Testing MCAR 

We can formally test MCAR, e.g. with a chi-squared test:

```{r, size="small"}
chisq.test(table(bpObs$ageCat, is.na(bpObs$bp))) 
```

Here we have strong evidence to reject MCAR.

## Missing at random

- BP ($Y_{2}$) is missing at random (MAR) given age ($Y_{1}$) if missingness is independent of BP ($Y_{2}$) given age ($Y_{1}$).
- This means that amongst subjects of the same age, missingness in BP is independent of BP.
- In terms of the missingness indicator $R$, MAR means

$$P(R=1|Y_{1},Y_{2})=P(R=1|Y_{1})$$

- The name is unfortunate. MAR does **not** mean data are missing completely randomly!

## Checking MAR

- We cannot check whether MAR holds based on the observed data.
- To do this we would need to check whether, within categories of age, those with missing BP had higher/lower BP than those with it observed.

## Blood pressure MAR given age

Using the full/complete data:

```{r, echo=FALSE, fig.height=5}
ggplot(bpFull, aes(x = ageCat, y = bp, color=rCat)) + 
  geom_jitter(width=0.05, height=0) +
  labs(x = "Age group", y="Systolic blood pressure (mmHg)") +
  theme(text = element_text(size=15))

```

From this MAR appears plausible - within age categories, the distributions of observed and missing BP look similar.

## Blood pressure MAR given age

But in reality all we get to see is:

```{r, echo=FALSE, fig.height=5, warning=FALSE}
ggplot(bpObs, aes(x = ageCat, y = bp)) + 
  geom_jitter(width=0.05, height=0) +
  labs(x = "Age group", y="Systolic blood pressure (mmHg)") +
  theme(text = element_text(size=15))
```

## Analysis assuming MAR

- If we are willing to assume data are MAR, we can construct unbiased estimates using a variety of statistical methods.
- e.g. estimate overall mean BP by a weighted average of observed BP means, weighting according to overall proportions of age categories:

```{r, echo=FALSE}
mean1 <- mean(bpObs$bp[bpObs$ageCat=="30-50 years"], na.rm=T)
mean2 <- mean(bpObs$bp[bpObs$ageCat=="50-70 years"], na.rm=T)
wgtavg <- 0.5*mean1 + 0.5*mean2
```

$$	\frac{100 \times `r round(mean1,1)` + 100 \times `r round(mean2,1)`}{200} = `r round(wgtavg,1)`  $$

- Note this is not the same as crude average observed BP.

```{r, size="small"}
mean(bpObs$bp, na.rm=TRUE)
```

## A different representation of MAR

- We have defined MCAR and MAR in terms of how $P(R=1|Y_{2},Y_{1})$ depends on age ($Y_{1}$) and BP ($Y_{2}$).
- From the plot, we see that MAR can also be viewed in terms of the conditional distribution of BP ($Y_{2}$) given age ($Y_{1}$).
- MAR implies that

$$	f(Y_{2}|Y_{1},R=0)=f(Y_{2}|Y_{1},R=1)=f(Y_{2}|Y_{1})$$

- That is, the distribution of BP ($Y_{2}$), given age ($Y_{1}$), is the same whether or not BP ($Y_{2}$) is observed.
- This key consequence of MAR is directly exploited by **multiple imputation**.


## Missing not at random

- If data are neither MCAR nor MAR, they are missing not at random (MNAR).
- This means the chance of seeing $Y_{2}$ depends on $Y_{2}$, even after conditioning on $Y_{1}$.
- Equivalently, $f(Y_{2}|Y_{1},R=0) \neq f(Y_{2}|Y_{1},R=1)$.
- MNAR is much more difficult to handle. Essentially the data cannot tell us how the missing values differ to the observed values (given $Y_{1}$).
- We are thus led to conducting sensitivity analyses.

## An MNAR analysis of mean blood pressure

- Suppose that, within age categories, the missing BPs are 10mmHg higher than the observed BPs.
- \textcolor{red}{Given} this assumption, we can estimate mean BP by assuming the mean of the missing BPs are 10mmHg higher than predicted by MAR:

```{r echo=FALSE}
mnarEst <- (sum(r[ageCat=="30-50 years"])*mean1 + sum(1-r[ageCat=="30-50 years"])*(mean1+10) + sum(r[ageCat=="50-70 years"])*mean2 + sum(1-r[ageCat=="50-70 years"])*(mean2+10))/n
```

$$ \frac{`r sum(r[ageCat=="30-50 years"])` \times `r round(mean1,1)` + `r sum(1-r[ageCat=="30-50 years"])` \times `r round(mean1+10,1)` + `r sum(r[ageCat=="50-70 years"])` \times `r round(mean2,1)` + `r sum(1-r[ageCat=="50-70 years"])` \times `r round(mean2+10,1)`}{`r n`} = `r round(mnarEst,1)`$$

- Note that we must specify how we think the missing BPs differ to the observed values, based on our contextual knowledge.
- The data **cannot** tell us how large this difference is!

## Summary

- Missing data introduce ambiguity into the analysis, beyond the familiar  sampling imprecision.
- Extra assumptions about the missingness mechanism are needed to ensure valid estimates and inferences.
- These assumptions can rarely be verified from the data at hand. 
- It is sensible to consider carefully possible missingness mechanisms, and formulate appropriate analyses.
- Because we cannot be sure about the type of missingness mechanism at work, sensitivity analyses are important.

## Summary continued

- Missingness mechanisms fall into three broad classes: MCAR, MAR and MNAR.
- Under MCAR, we obtain valid estimates and inferences by analysing the subset of subjects with no missing values.
- Under MAR, we must allow for variables (somehow) which predict missingness.
- MAR analyses can be done in a number of ways.
- Multiple imputation is one such approach, which we will explore in this course.

## References {.allowframebreaks}